Detection and replication of epistasis influencing transcription in humans

### 2013-08-31 docs/manuscript/nature_letter.tex
manuscript has been submitted to nature
more stuff on signs

\clearpage
\section{Tables}

@@

\clearpage
\section{Figures}

\begin{figure}[H]
	\centering
	\includegraphics[width=2.5in]{gpbonfrep.pdf}
	\caption{\textbf{Replication of GP maps in two independent populations} The GP maps for each epistatic interaction that is significant at the Bonferroni level in both replication datasets are shown. Each GP map consists of nine tiles where each tile represents the expression level for that two-locus genotype class. Phenotypes are for gene transcript levels (dark coloured tiles = high expression, light coloured tiles = low expression). Columns of GP maps are for each independent dataset. Rows of GP maps are for each of 30 significantly replicated interactions at the Bonferroni level, corresponding to the rows in Table \ref{tab:bonferroni}. There is a clear trend of the GP maps replicating across all three datasets.}
	\label{fig:gpmaps}
\end{figure}
\clearpage

\begin{figure}[H]
	\centering
	\includegraphics[width=5in]{pale_gray_graph_of_interactions_2_lists_GrayEdges}
	\caption{\textbf{Discovery and replication of epistatic networks} All 434 putative genetic interactions (edges) with data common to discovery and replication sets is shown, where black nodes represent SNPs and red nodes represent traits (gene expression probes). Three hundred and forty-five interactions had $p$-values exceeding the 2.5\% confidence interval following meta analysis of the replication data The remaining 89 interactions that did not replicate are depicted in grey. It is evident that a large proportion of the complex networks identified in the discovery set also exist in independent populations.}
	\label{fig:fireworks}
\end{figure}
\clearpage

\begin{figure}[H]
	\centering
	\includegraphics[width=2.5in]{qqMeta.pdf}
	\caption{\textbf{Q-Q plots of interaction $p$-values from replication datasets} The top panel shows all 434 discovery SNPs that were tested for interactions. Observed $p$-values ($y$-axis, $-\log_{10}$ scale) are plotted against the expected $p$-values ($x$-axis, $-\log_{10}$ scale). The multiple testing correction threshold for significance following Bonferroni correction is denoted by a dotted line. The bottom panel shows the same data as the top panel but excluding the 30 interactions that were significant at the Bonferroni level in the replication datasets. The shaded grey area represents the 5\% confidence interval for the expected distribution of $p$-values. Dark blue points represent $p$-values that exceed the confidence interval, light blue are within the confidence interval.}
	\label{fig:qqMeta}
\end{figure}
\clearpage

\clearpage
\section{Supplementary Figures}
\setcounter{figure}{0}
\makeatletter 
\renewcommand{\thefigure}{S\@arabic\c@figure} 
\makeatletter

@@

\clearpage
\section{Supplementary Tables}
\setcounter{table}{0}
\makeatletter 
\renewcommand{\thetable}{S\@arabic\c@table} 
\makeatletter

@@

\clearpage
% \section{References}
\bibliography{refs}

>>>>

\clearpage
\section*{Tables}

@@

\clearpage
\section*{Figures}

\begin{figure}[H]
	\centering
	% \includegraphics[width=2.5in]{gpbonfrep.pdf}
	\caption{\textbf{Replication of GP maps in two independent populations} The GP maps for each epistatic interaction that is significant at the Bonferroni level in both replication datasets are shown. Each GP map consists of nine tiles where each tile represents the expression level for that two-locus genotype class. Phenotypes are for gene transcript levels (dark coloured tiles = high expression, light coloured tiles = low expression). Columns of GP maps are for each independent dataset. Rows of GP maps are for each of 30 significantly replicated interactions at the Bonferroni level, corresponding to the rows in Table \ref{tab:bonferroni}. There is a clear trend of the GP maps replicating across all three datasets.}
	\label{fig:gpmaps}
\end{figure}

\begin{figure}[H]
	\centering
	% \includegraphics[width=2.5in]{qqMeta.pdf}
	\caption{\textbf{Q-Q plots of interaction $p$-values from replication datasets} The top panel shows all 434 discovery SNPs that were tested for interactions. Observed $p$-values ($y$-axis, $-\log_{10}$ scale) are plotted against the expected $p$-values ($x$-axis, $-\log_{10}$ scale). The multiple testing correction threshold for significance following Bonferroni correction is denoted by a dotted line. The bottom panel shows the same data as the top panel but excluding the 30 interactions that were significant at the Bonferroni level in the replication datasets. The shaded grey area represents the 5\% confidence interval for the expected distribution of $p$-values. Dark blue points represent $p$-values that exceed the confidence interval, light blue are within the confidence interval.}
	\label{fig:qqMeta}
\end{figure}

\begin{figure}[H]
	\centering
	% \includegraphics[width=5in]{pale_gray_graph_of_interactions_2_lists_GrayEdges}
	\caption{\textbf{Discovery and replication of epistatic networks} All 434 putative genetic interactions (edges) with data common to discovery and replication sets is shown, where black nodes represent SNPs and red nodes represent traits (gene expression probes). Three hundred and forty-five interactions had $p$-values exceeding the 2.5\% confidence interval following meta analysis of the replication data The remaining 89 interactions that did not replicate are depicted in grey. It is evident that a large proportion of the complex networks identified in the discovery set also exist in independent populations.}
	\label{fig:fireworks}
\end{figure}

\bibliography{refs}
\clearpage

\section*{Supplementary Figures}
\setcounter{figure}{0}
\makeatletter 
\renewcommand{\thefigure}{S\@arabic\c@figure} 
\makeatletter

@@

\clearpage
\section*{Supplementary Tables}
\setcounter{table}{0}
\makeatletter 
\renewcommand{\thetable}{S\@arabic\c@table} 
\makeatletter

@@

\begin{table}[ht]
\begin{threeparttable}
\caption{\textbf{Concordance of sign of epistatic variance components between discovery and replication datasets}}
\label{tab:sign}
\begin{tabular}{lllrrrr}

  \hline
Test & Interactions & Dataset & $n$ & Expected & Observed & $p$-value \\
  \hline
1\tnote{a} & All & EGCUT & 434 & 217.00 & 306 & $6.69 \times 10^{18}$ \\
&  & Fehrmann & 434 & 217.00 & 278 & $5.04 \times 10^{09}$ \\
&  & Both & 434 & 108.50 & 221 & $5.56 \times 10^{31}$ \\
& Significant & EGCUT & 30 & 15.00 & 25 & $3.25 \times 10^{04}$ \\
&  & Fehrmann & 30 & 15.00 & 24 & $1.43 \times 10^{03}$ \\
&  & Both & 30 & 7.50 & 22 & $3.76 \times 10^{08}$ \\
  \hline
2\tnote{b} & All & EGCUT & 434 & 54.25 & 92 & $4.22 \times 10^{07}$ \\
&  & Fehrmann & 434 & 54.25 & 79 & $6.18 \times 10^{04}$ \\
&  & Both & 434 & 6.78 & 30 & $2.55 \times 10^{11}$ \\
& Significant & EGCUT & 30 & 3.75 & 19 & $9.46 \times 10^{11}$ \\
&  & Fehrmann & 30 & 3.75 & 19 & $9.46 \times 10^{11}$ \\
&  & Both & 30 & 0.47 & 18 & $2.23 \times 10^{25}$ \\
  \hline
3\tnote{c} & All & EGCUT & 434 & 27.12 & 34 & $1.65 \times 10^{01}$ \\
&  & Fehrmann & 434 & 27.12 & 35 & $1.35 \times 10^{01}$ \\
&  & Both & 434 & 1.70 &  2 & $6.89 \times 10^{01}$ \\
& Significant & EGCUT & 30 & 1.88 &  8 & $3.92 \times 10^{04}$ \\
&  & Fehrmann & 30 & 1.88 &  9 & $6.22 \times 10^{05}$ \\
&  & Both & 30 & 0.12 &  1 & $1.11 \times 10^{01}$ \\
  \hline
4\tnote{d} & All & EGCUT & 1133 & 566.50 & 775 & $7.10 \times 10^{36}$ \\
&  & Fehrmann & 1133 & 566.50 & 726 & $1.90 \times 10^{21}$ \\
&  & Both & 1133 & 283.25 & 562 & $1.39 \times 10^{70}$ \\
& Significant & EGCUT & 73 & 36.50 & 55 & $1.69 \times 10^{05}$ \\
&  & Fehrmann & 73 & 36.50 & 55 & $1.69 \times 10^{05}$ \\
&  & Both & 73 & 18.25 & 46 & $7.86 \times 10^{12}$ \\
   \hline
\end{tabular}

\begin{tablenotes}
\item[a] The sign of the most significant epistatic variance component in discovery is the same as the corresponding variance component in the replication data.
\item[b] The largest epistatic variance component in the discovery is the same as in the replication with the same sign in both.
\item[c] The sign of all four epistatic variance components are identical in the discovery and the replication.
\item[d] The sign of all epistatic variance components in the discovery with $p < 0.05$ are the same as the corresponding variance components in the replication data.
\end{tablenotes}

\end{threeparttable}
\end{table}
